---
title: "Short Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ocf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(ocf)
```

In this tutorial, we show how to use the `ocf` package to estimate and make inference about the conditional choice probabilities and the covariates' marginal effects. 

For illustration purposes, we use the synthetic data set provided in the `orf` package:

```{r data-generation, eval = FALSE}
## Load data from orf package.
set.seed(1986)

library(orf)
data(odata)

y <- as.numeric(odata[, 1])
X <- as.matrix(odata[, -1])
```

### Conditional Probabilities
The `ocf` function constructs a collection of forests, one for each category of `y` (three in this case). We can then use the forests to predict out-of-sample conditional probabilities using the `predict` method. By default, `predict` returns a matrix with the predicted probabilities and a vector of predicted class labels (each observation is labelled to the highest-probability class). 

```{r adaptive-ocf, eval = FALSE}
## Training-test split.
train_idx <- sample(seq_len(length(y)), floor(length(y) * 0.5))

y_tr <- y[train_idx]
X_tr <- X[train_idx, ]

y_test <- y[-train_idx]
X_test <- X[-train_idx, ]

## Fit ocf on training sample. Use default settings.
forests <- ocf(y_tr, X_tr)

## Summary of data and tuning parameters.
summary(forests)

## Out-of-sample predictions.
predictions <- predict(forests, X_test)

head(predictions$probabilities)
table(y_test, predictions$classification)
```

We can also implement honesty, which is a necessary condition to produce asymptotically normal and consistent predictions. In the following, we set `honesty = TRUE` to construct honest forests.

```{r honest-ocf, eval = FALSE}
## Honest forests.
honest_forests <- ocf(y_tr, X_tr, honesty = TRUE)
honest_predictions <- predict(honest_forests, X_test)

## Compare predictions with adaptive fit.
cbind(head(predictions$probabilities), head(honest_predictions$probabilities))
```

To estimate standard errors for the predicted probabilities, we set `inference = TRUE`. This requires also to set `honesty = TRUE`: the formula for the variance is valid only for honest predictions. The estimation of standard errors considerably slows down the routine. However, we can increase the number of threads used to construct the forests to speed up the routine.

```{r honest-ocf-inference, eval = FALSE}
## Compute standard errors.
honest_forests <- ocf(y_tr, X_tr, honesty = TRUE, inference = TRUE, n.threads = 0) # Use all CPUs.
head(honest_forests$predictions$standard.errors)
```

### Covariates' Marginal Effects
The `marginal_effects` function post-processes the predictions to estimate mean marginal effects, marginal effects at the mean, or marginal effects at the median, according to the `eval` argument. In the following, we construct our forests in the training sample and use them to estimate the marginal effects at the mean in the test sample.

```{r adaptive-me, eval = FALSE}
## Fit ocf on training sample.
forests <- ocf(y_tr, X_tr)

## Marginal effects at the mean on test sample.
me_atmean <- marginal_effects(forests, data = X_test, eval = "atmean")
summary(me_atmean)
```

As before, we can set `inference = TRUE` to estimate the standard errors. Again, this requires the use of honest forests and considerably slows down the routine.

```{r honest-me, eval = FALSE}
## Honest forests.
honest_forests <- ocf(y_tr, X_tr, honesty = TRUE) # Notice we do not need inference here!

## Compute standard errors.
honest_me_atmean <- marginal_effects(honest_forests, data = X_test , eval = "atmean", inference = TRUE)

## LATEX.
print(honest_me_atmean, latex = TRUE)
```
